---
title: "Mexico City Industrial Competitiveness"
subtitle: "MIT REAP: Mexico City Focuse"
author: "MDS. René Rosado González\nTecnologico de Monterrey"
output: 
  html_document:
    theme: paper
    highlight: haddock
---

```{r setup, include = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("~/Desktop/ITESM/MIT REAP/")

# Instal - load tidyverse                                                       
if(require(tidyverse) == FALSE){                                                
  install.packages('tidyverse')                                                 
  library(tidyverse)                                                            
}else{                                                                          
  library(tidyverse)                                                            
}

# Instal - load plotly                                                
if(require(plotly) == FALSE){                                                
  install.packages('plotly')                                                 
  library(plotly)                                                            
}else{                                                                          
  library(plotly)                                                            
}

#  Instal - DT                                                     
if(require(DT) == FALSE){                                                
  install.packages('DT')                                                 
}  

# Instal - load rvest                                                       
if(require(rvest) == FALSE){                                                
  install.packages('rvest')                                                 
  library(rvest)                                                            
}else{                                                                          
  library(rvest)                                                            
}         

# Instal - load httr                                                       
if(require(httr) == FALSE){                                                
  install.packages('httr')                                                 
  library(httr)                                                            
}else{                                                                          
  library(httr)                                                            
}    

# Instal  rjson                                                       
if(require(rjson) == FALSE){                                                
  install.packages('rjson')                                                 
}
```

```{r get_data, include = FALSE, warning=FALSE}
# Load Economic Census
mexico_city_economic_census = read_csv("data/raw/ce2019_cdmx.csv") %>% 
  # Clean names
  janitor::clean_names() %>% 
  # Keep industry level and non firms size strata
  filter(nchar(codigo) == 6, is.na(id_estrato)) %>% 
  # Get ids
  mutate(
    entidad = as.numeric(entidad),
    sector_id = str_extract(codigo,'\\d{2}'),
    subsector_id = str_extract(codigo,'\\d{3}'),
    industry_group_id = str_extract(codigo,'\\d{4}'),
    naics_industry_id = str_extract(codigo,'\\d{5}')) %>%
  # Select relevant variables
  select(
    state_id = entidad, 
    municipality_id = municipio,
    national_industry_id = codigo, 
    sector_id:industry_group_id,
    economic_units = ue, 
    gross_production = a111a, 
    gross_added_value = a131a,
    total_expenses = a700a,
    total_income = a800a,
    employees = h001a,
    worked_hours = h001d)  %>% 
  # Create new variables
  mutate(gav_per_worked_hour = 1000 * gross_added_value/worked_hours,
         productivity = 1000 * gross_production/worked_hours) %>% 
  # Keep only city aggregates
  subset(is.na(municipality_id)) %>% 
  # Get rid of the municipality column
  mutate(municipality_id = NULL)

# Get data from Data Mexico
# Economic Complexity Index (ECI) by municipality
economic_complexity_index_municipality = GET(
  # Set GET request
  url = "https://datamexico.org/api/data?Quarter=20214&cube=complexity_eci&drilldowns=ECI+Ranking,Municipality&measures=ECI&parents=false&sparse=false&locale=es") %>% 
  # Read the request and strucutre de json data into tidyformat
  read_html() %>% html_text() %>% rjson::fromJSON() %>% .[['data']] %>%
  # Bind rows into a tibble and clean names
  bind_rows() %>% janitor::clean_names() %>% 
  # Rename the ECI column
  rename(economic_complexity_index = eci)

# Economic Activities Complexity Index (ACI)
activities_complexity_index = GET(
  # Set GET request
  url = "https://datamexico.org/",
  path = "api/stats/pci",
  query = list(
    cube="inegi_denue",
    Month="20211117,20210510,20201119",
    rca="State,National Industry,Number of Employees Midpoint",
    threshold="National Industry:900,State:900",
    locale="en"
  )) %>% 
  # Read the request and strucutre de json data into tidyformat
  read_html() %>% html_text() %>% rjson::fromJSON() %>% .[['data']] %>%
  # Bind rows into a tibble and clean names
  bind_rows() %>% janitor::clean_names() %>% 
  # Rename the ACI column
  rename(activities_complexity_index = number_of_employees_midpoint_pci)


# Mexico City Relative Comparative Advantage (RCA) - Data Mexico
relative_comparative_advantage_cdmx = GET(
  # Set GET request
  url = "https://datamexico.org/",
  path = "api/stats/relatedness",
  query = list(
    cube="inegi_denue",
    Month="20211117,20210510,20201119",
    rca="State,National Industry,Number of Employees Midpoint",
    parents="true",
    filter_State="9",
    threshold="National Industry:900,State:900",
    locale="en"
  )) %>% 
  # Read the request and strucutre de json data into tidyformat
  read_html() %>% html_text() %>% rjson::fromJSON() %>% .[['data']] %>%
  # Bind rows into a tibble and clean names
  bind_rows() %>% janitor::clean_names() %>% 
  # Rename relevant columns
  rename(employees_number = number_of_employees_midpoint,
         affinity = number_of_employees_midpoint_relatedness,
         relative_comparative_advantage = number_of_employees_midpoint_rca)

# Join all data
industry_advantages = activities_complexity_index %>% 
  inner_join(relative_comparative_advantage_cdmx, by = c("national_industry_id", "national_industry")) %>% 
  left_join(mexico_city_economic_census, by = c("national_industry_id", "state_id", "sector_id", 
                               "subsector_id", "industry_group_id")) %>% 
  # Clean and order date
  mutate(state_id = paste0('0',state_id)) %>% 
  select(state_id, state, sector_id, sector, subsector_id, subsector, 
         industry_group, industry_group_id, naics_industry_id, naics_industry,
         national_industry_id, national_industry, employees_number,
         activities_complexity_index, affinity, relative_comparative_advantage,
         gross_production:total_income,worked_hours:productivity
         ) %>% 
  mutate_if(is.numeric,round,2) %>% 
  # Order by Relative Competitive Advantage
  arrange(desc(relative_comparative_advantage))
```


### Metrics for the assesment

This report analyze which economic activities have the highest potential to boost innovation and entrepreneurial development in Mexico City. In order to do so, the three following metrics are taken into account.

1. **Economic Complexity** uses data on the geography of economic activities to infer the presence of capabilities. This capabilities are summarized in the **The Economic Complexity Index (ECI)**. 

2. **Affinity (relatedness)** measures the probability of a city, or region to adopt a certain economic activity (a new product, industry, technology). According to this principle, the probabilities of a city developing a new industry increases with the number of related activities present in that location. This principle can be used to predict which activities are most likely to emerge - or disappear - in a city.

3. **Revealed Comparative Advantage (RCA)** measures the degree of specialization of a certain region. The RCA is equal to the proportion of exports of a certain product, divided by the proportion of world exports of this product. If the RCA value of a certain activity is higher than 1, then the territory is a specialist in its production.

The following table presents the economic activities avaiblable in Mexico City. They are ordered from the highest to the lowest RCA. 

```{r table, echo = FALSE, warning = FALSE}
DT::datatable(industry_advantages, 
              colnames=c("State ID", "State", "Sector ID", "Sector", "Subsector ID", "Subsector", "Industry Group ID", "Industry Group", "NAICS ID", "NAICS", "Industry ID", "Industry", "Midpoint Employees", "Activity Complexity (ACI)", "Affinity (Relatedness)", "Relative Comparative Advantege (RCA)", "Gross Production (Million MXN)", "Gross Added Value (GAV in million MXN)", "Total Expenses (Million MXN)", "Total Income (Million MXN)", "Worked Hours (Thousands)","GAV per Worked Hour (MXN)", "Productivity (MXN)"),
              width = '100%', 
              options = list(scrollX = TRUE))
```

Here are some visualizations to facilitate your analysis:

```{r plot_1, echo = FALSE, warning = FALSE, out.width="100%", out.height="100%"}

hline = function(y = 1) {
  list(
    type = "line",
    x0 = 0,
    x1 = 1,
    xref = "paper",
    y0 = y,
    y1 = y,
    line = list(color = "gray50", dash="dot", with = 0.75)
  )
}

plot_ly(data = industry_advantages) %>% 
  add_trace( 
    type = "scatter", 
    mode = "markers",
    x = ~activities_complexity_index,
    y = ~relative_comparative_advantage, 
    hoverinfo = list(type='text'),
    text = ~paste(national_industry,
                  "<br>Midpoint Employees: ", employees_number, 
                  '<br>Comparative Advantage:', round(relative_comparative_advantage,2),
                  '<br>Gross Production: ', round(gross_production,2),
                  '<br>Gross Added Value: ', round(gross_added_value,2),
                  '<br>GAV per Worked Hour: ', round(gav_per_worked_hour,2)),
    color = ~affinity, 
    colors='Reds',
    showlegend = FALSE,
    size=~employees_number,
    sizes = c(10, 50),
    showlegend = F,
    marker = list(
      line = list(color = "black",width = 1), 
      sizemode = 'diameter',showscale = FALSE
      )
    #width = 1280, height = 800, 
    ) %>% 
  layout(
    title = list(text = "Mexico City Industrial Complexity"), 
    font  = list(family = "Montserrat", size = 12, color = "gray10"), 
    xaxis = list(title = list(text ='Economic Activities Complexity'),
                 showgrid = FALSE),
    yaxis = list(title = list(text ='Comparative Advantage'),
                 zeroline = FALSE, showgrid = FALSE),
    shapes = list(hline(1)),
    margin = list(l = 50,r = 50,b = 100,t = 100, pad = 5)
         ) %>% 
  colorbar(title = "Affinity\n(relatedness)",
           tickfont = list(size = 12),
           line = list(color = 'white'))

```

```{r plot_2, echo = FALSE, warning = FALSE, out.width="100%", out.height="100%"}

economic_complexity_index_municipality %>% 
  head(200) %>% 
  mutate(cdmx = if_else(str_detect(municipality_id,'^9'),'CDMX','Other')) %>% 
  plot_ly() %>% 
  add_trace( 
    type = "bar", 
    x = ~eci_ranking,
    y = ~economic_complexity_index, 
    hoverinfo = list(type='text'),
    text = ~paste(municipality,
                  "<br>Municipality ID: ", municipality_id, 
                  '<br>ECI:', round(economic_complexity_index,2)
                  ),
    color = ~cdmx
    ) %>% 
  layout(
    title = list(text = "Top 200 - Most Complex Mexican Municipalities"), 
    font  = list(family = "Montserrat", size = 12, color = "gray10"), 
    xaxis = list(title = list(text ='Municipalities'),
                 showgrid = FALSE),
    yaxis = list(title = list(text ='ECI'),
                 zeroline = FALSE, showgrid = FALSE),
    margin = list(l = 50,r = 50,b = 100,t = 100, pad = 5)
         )

```
